\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{titlesec}
\usepackage[
  backend=biber,
  style=numeric,
  sorting=none
]{biblatex}
\addbibresource{references.bib}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tabularx}

\titleformat{\section}{\large\bfseries}{\thesection}{0.5em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.5em}{}

\begin{document}

%==============================================================================
% Title block
%==============================================================================
\begin{center}
  {\Large \textbf{Classification and Measurement of One-Dimensional Hyperuniform Point Patterns}}\\[2ex]
  Michael Fang \\
  \today
\end{center}

\noindent
{\small
\begin{minipage}{\textwidth}
\centering

\begin{minipage}[t]{0.44\textwidth}\vspace{0pt}
Primary advisor: Professor Salvatore Torquato\\
Second reader: TBD
\end{minipage}
\hspace{0.04\textwidth}%
\begin{minipage}[t]{0.40\textwidth}\vspace{0pt}
\raggedleft
This paper represents my own work in accordance with University regulations.\\
/s/\ Michael Fang
\end{minipage}

\end{minipage}
}
\vspace{-0.2cm}

%==============================================================================
% Abstract
%==============================================================================
\begin{abstract}
We present a systematic numerical study of hyperuniformity in one-dimensional point
patterns, characterizing 27 systems that span all three universality classes.
Number variance, structure factor, and diffusion spreadability serve as complementary
probes of long-wavelength density fluctuations, validated against exact analytical
benchmarks before application to the full pattern catalog. We show that all five
metallic-mean quasicrystal chains (Fibonacci through Nickel) are constrained to the
hyperuniformity exponent $\alpha = 3$ by the unimodular determinant condition on their
$2\!\times\!2$ substitution matrices, and report surface-area coefficients $\bar\Lambda$
for all five chains, four of which are novel. Moving to $3\!\times\!3$ matrices, we
identify the Bombieri-Taylor cubic substitution as the first known one-dimensional
quasicrystal with $1 < \alpha < 2$, obtaining $\alpha = 1.545$ in quantitative agreement
with the eigenvalue formula of Oğuz et al. An exhaustive enumeration of 39,304 candidate
$3\!\times\!3$ non-negative integer matrices proves algebraically that the interval
$2 < \alpha < 3$ is inaccessible at this matrix rank, while yielding a new deterministic
substitution tiling with $\alpha = 2$ exactly --- to our knowledge the first quasiperiodic
structural analogue of the stochastic uncorrelated random lattice. Generalized scaling
coefficients are introduced to rank Class~II and~III patterns, and stealthy hyperuniform
configurations are found to exist in two distinct structural phases that both satisfy the
stealthy constraint while differing substantially in short-range order.
\end{abstract}

\tableofcontents
\clearpage

%==============================================================================
\section{Introduction}
\label{sec:introduction}
%==============================================================================

Long-range density fluctuations are a central diagnostic of order in many-body systems.
When points are distributed at random with no correlations, the variance of the count
inside a window of half-width $R$ grows in proportion to the window length:
$\sigma^2_\text{Poisson}(R) = 2\rho R$, where $\rho$ is the mean density. This linear
growth is the Poisson baseline --- the generic behavior of uncorrelated disordered matter.
\textit{Hyperuniform} systems are those for which large-scale density fluctuations are
anomalously suppressed relative to this baseline~\cite{torquato2003local,torquato2018hyperuniform}.
Hyperuniformity is not a statement about short-range structure: a hyperuniform pattern can
appear locally disordered while being globally organized in a precise statistical sense.
This combination of apparent local randomness with long-range suppression of density
fluctuations places hyperuniform systems in a structurally distinct class, intermediate
between perfect crystals and uncorrelated fluids.

Hyperuniform arrangements appear across a remarkable range of physical and biological
contexts. The photoreceptor mosaic of bird retinas forms a hyperuniform pattern, a design
that enables near-uniform visual sampling without the periodic artifacts of a
crystal~\cite{torquato2018hyperuniform}. Systems of hard spheres near their jamming point
exhibit hyperuniformity, suggesting that the suppression of density fluctuations is a
signature of mechanical criticality. At the largest accessible scales, the distribution of
galaxies in cosmological surveys appears nearly hyperuniform, hinting at long-range
correlations established in the early universe. Quasicrystals --- aperiodic materials
with long-range orientational order but no translational periodicity, first observed in
an aluminium-manganese alloy by Shechtman and coworkers~\cite{shechtman1984metallic} ---
are naturally hyperuniform because their long-range aperiodic structure prevents macroscopic
density imbalances. From an engineering perspective, computationally designed disordered
hyperuniform networks have been shown to support photonic band gaps rivaling those of
periodic photonic crystals, offering a route to isotropic wave-guiding that does not
require strict periodicity~\cite{torquato2018hyperuniform}.

The formal definition of hyperuniformity, due to Torquato and
Stillinger~\cite{torquato2003local}, is that the structure factor $S(k)$ --- the
Fourier-space measure of density fluctuations at spatial wavelength $2\pi/k$ --- satisfies
$S(k) \to 0$ as $k \to 0$. The \textit{rate} of this vanishing defines a spectrum of
hyperuniformity. For systems where $S(k) \sim |k|^\alpha$ near $k = 0$, the exponent
$\alpha > 0$ quantifies the degree of suppression: larger $\alpha$ means faster vanishing
and more strongly reduced fluctuations at long wavelengths. Three universality classes arise
from the asymptotic behavior of $\sigma^2(R)$~\cite{torquato2018hyperuniform}. Class~I
($\alpha > 1$) is the strongest: the number variance is bounded at large $R$, approaching
a constant plateau. Perfect crystals and many quasicrystals belong to this class, as do
stealthy hyperuniform systems for which $S(k) = 0$ over an entire exclusion zone near
$k = 0$. Class~II ($\alpha = 1$) is the marginal case where $\sigma^2(R)$ grows
logarithmically; the period-doubling substitution chain is a one-dimensional example.
Class~III ($0 < \alpha < 1$) gives power-law variance growth strictly slower than the
Poisson baseline; certain substitution chains with non-Pisot eigenvalues fall here.

One spatial dimension is a natural test bed for hyperuniformity theory. Point patterns in
1D can be generated deterministically to sizes $N \sim 10^7$ by iterating substitution
rules, exact analytical benchmarks exist, and the connection between a pattern's algebraic
structure and its hyperuniformity class is particularly transparent. Yet the 1D landscape
of known Class~I patterns is sparse. The survey in Ref.~\cite{torquato2018hyperuniform}
lists only a handful of examples, and the interval $1 < \alpha < 2$ was previously
unoccupied by any known 1D quasicrystal. As we show below, the reason is algebraic: all
$2\times 2$ metallic-mean substitution matrices are unimodular ($|\det M| = \pm 1$), which forces
$\alpha = 3$ exactly~\cite{oguz2019substitution}. Filling the $1 < \alpha < 2$ gap
requires moving outside this family.

In this work, we characterize 27 one-dimensional point patterns spanning all three
hyperuniformity classes. The patterns include the five metallic-mean quasicrystals
(Fibonacci, Silver, Bronze, Copper, and Nickel chains), all of which have $\alpha = 3$ by
the determinant constraint; uncorrelated random lattice (URL) perturbations, which realize
$\alpha = 2$ exactly and serve as key validation benchmarks; stealthy hyperuniform
configurations generated by collective coordinate optimization; and several substitution
chains that realize Class~II and Class~III behavior. The central new result is the
identification of the Bombieri-Taylor cubic substitution~\cite{bombieri1986which} as a
Class~I quasicrystal with $\alpha = 1.545$, the first known example with
$1 < \alpha < 2$~\cite{oguz2019substitution}. We also report surface-area coefficients
$\bar{\Lambda}$ for the Silver, Bronze, Copper ($\bar{\Lambda} = 0.293$), and Nickel
($\bar{\Lambda} = 0.310$) chains; the Copper and Nickel values are reported here for
the first time.

The remainder of this paper is organized as follows.
Section~\ref{sec:theory} develops the theoretical framework: the number variance and its
relation to the pair correlation function, the structure factor and hyperuniformity exponent,
the surface-area coefficient $\bar{\Lambda}$, and the eigenvalue formula for substitution
tilings.
Section~\ref{sec:methods} describes the numerical methods and validation benchmarks used
throughout this work. Results and conclusions will follow in subsequent sections.

%==============================================================================
\section{Theoretical Background}
\label{sec:theory}
%==============================================================================

\subsection{Number Variance}
\label{sec:variance}

To quantify how density fluctuations depend on scale, we examine how many points fall
inside a window of half-width $R$ placed at a random position along a one-dimensional
point pattern of mean linear density $\rho$. The variance of this count across window
placements is the \textit{number variance} $\sigma^2(R)$. For a Poisson (completely
uncorrelated) point process with density $\rho$, the variance grows linearly with window
size: $\sigma^2_\text{Poisson}(R) = 2\rho R$~\cite{torquato2003local}. This Poisson
scaling is the baseline against which all other patterns are measured --- it reflects the
fact that in an uncorrelated system, each region of space is independent, so larger windows
simply accumulate more independent uncertainty.

For a general stationary ergodic point process, $\sigma^2(R)$ can be expressed in terms of
the pair correlation function $g_2(r)$, which encodes the probability of finding a second
point at distance $r$ from a reference point relative to the uncorrelated
expectation~\cite{torquato2003local}:
\begin{equation}
  \sigma^2(R) = 2\rho R \left[ 1 + 2\rho \int_0^{2R} \!\left(1 - \frac{r}{2R}\right)
  \bigl[g_2(r) - 1\bigr]\, dr \right].
  \label{eq:variance-integral}
\end{equation}
The first term, $2\rho R$, is the uncorrelated (Poisson) contribution. The integral
represents the correction from inter-point correlations: where $g_2(r) > 1$ (clustering),
the variance increases; where $g_2(r) < 1$ (anticorrelation), the variance decreases.
A pattern is \textbf{hyperuniform} if the integral correction exactly cancels the Poisson
term in the large-$R$ limit~\cite{torquato2003local}:
\begin{equation}
  \lim_{R \to \infty} \frac{\sigma^2(R)}{2\rho R} = 0.
  \label{eq:hyperuniform-def}
\end{equation}
Physically, this requires that points be anticorrelated at large separations in precisely
the right amount: $\int_0^\infty [g_2(r)-1]\,dr = -1/(2\rho)$.

Based on how $\sigma^2(R)$ grows asymptotically, hyperuniform patterns fall into three
universality classes~\cite{torquato2018hyperuniform}:
\begin{itemize}
  \item \textbf{Class~I} ($\alpha > 1$): $\sigma^2(R)$ approaches a finite (though
        possibly oscillating) constant at large $R$. These patterns exhibit the strongest
        suppression of density fluctuations. Perfect crystals and quasicrystals generated
        by Pisot substitution rules belong here.
  \item \textbf{Class~II} ($\alpha = 1$): $\sigma^2(R) \sim C\ln R$. The variance grows
        without bound, but only logarithmically. The period-doubling chain is a 1D example.
  \item \textbf{Class~III} ($0 < \alpha < 1$): $\sigma^2(R) \sim R^{1-\alpha}$, a
        power-law growth strictly slower than the Poisson baseline. Aperiodic chains with
        non-Pisot eigenvalue spectra can realize this behavior.
\end{itemize}
Non-hyperuniform patterns ($\alpha \leq 0$) have $\sigma^2(R) \sim R$ or faster.

\subsection{Structure Factor and Hyperuniformity Exponent}
\label{sec:alpha}

The same information encoded in $\sigma^2(R)$ is also captured by the \textit{structure
factor} in Fourier space. For a point pattern with $N$ points at positions $\{x_j\}$,
\begin{equation}
  S(k) = \frac{1}{N}\left|\sum_{j=1}^{N} e^{ikx_j}\right|^2.
  \label{eq:sk-def}
\end{equation}
Physically, $S(k)$ measures the intensity of density fluctuations at spatial wavelength
$2\pi/k$. For a Poisson process $S(k) = 1$ at all $k$; for a perfect crystal $S(k) = 0$
for all $|k|$ below the first reciprocal lattice vector. The two observables are related
by an exact integral~\cite{torquato2003local}:
\begin{equation}
  \sigma^2(R) = \frac{\rho}{\pi} \int_0^\infty S(k)\,
  \left(\frac{2\sin(kR)}{k}\right)^2 dk.
  \label{eq:variance-sk}
\end{equation}
Hyperuniformity requires $S(k) \to 0$ as $k \to 0$, which causes the integrand to vanish
faster than for Poisson and yields sub-linear variance growth at large $R$.

For systems where $S(k)$ approaches zero as a power law,
\begin{equation}
  S(k) \sim B|k|^\alpha \quad \text{as } k \to 0,
  \label{eq:sk-power-law}
\end{equation}
the exponent $\alpha > 0$ is the \textit{hyperuniformity exponent}. Larger $\alpha$ implies
more strongly suppressed fluctuations at long wavelengths. For quasicrystals, the structure
factor consists of dense Bragg peaks rather than a smooth continuous envelope, so the
exponent is extracted from the integrated intensity $Z(k) = 2\int_0^k S(q)\,dq \sim
k^{1+\alpha}$ near $k = 0$~\cite{oguz2019substitution}. In practice, $\alpha$ is also
estimated from the asymptotic slope of $\sigma^2(R)$ on a log-log scale, providing an
independent numerical check.

\subsection{Surface-Area Coefficient $\bar{\Lambda}$}
\label{sec:lambda}

Within Class~I, all patterns have bounded number variance, but two systems with the same
$\alpha$ can differ substantially in their residual local disorder. The \textit{surface-area
coefficient} $\bar{\Lambda}$ is the secondary metric that resolves this ambiguity and
provides an absolute ranking within Class~I.

For Class~I systems in one dimension, $\sigma^2(R)$ does not simply converge to a constant
but typically oscillates about a plateau, particularly for quasicrystals with dense Bragg
spectra. A pointwise limit is therefore not well-defined in general. Instead, $\bar{\Lambda}$
is defined as the long-range running average~\cite{torquato2018hyperuniform,torquato2003local}:
\begin{equation}
  \bar{\Lambda} = \lim_{R \to \infty} \frac{1}{R}\int_0^R \sigma^2(R')\, dR'.
  \label{eq:lambda-def}
\end{equation}
This running average converges to the mean level of the oscillating variance plateau, and
reduces to the ordinary limit $\lim_{R\to\infty} \sigma^2(R)$ whenever that limit exists.
For systems where the variance oscillates without converging, Eq.~\eqref{eq:lambda-def}
extracts a unique, well-defined amplitude.

The name ``surface-area coefficient'' refers to the $d$-dimensional interpretation: for a
Class~I pattern in $d$ dimensions, the variance scales at large $R$ as
$\sigma^2(R) \sim 2\bar{\Lambda}\, R^{d-1}$, where $R^{d-1}$ is proportional to the
surface area of the observation window. In $d=1$ this reduces to $\sigma^2(R) \to \bar\Lambda$
(a constant), consistent with the definition Eq.~\eqref{eq:lambda-def}.
The quantity $\bar{\Lambda}$ is the primary ranking metric within Class~I: the perfect
integer lattice achieves the global minimum $\bar{\Lambda} = 1/6$, proved exact in
Ref.~\cite{torquato2003local}, and any deviation from perfect periodicity strictly increases
$\bar{\Lambda}$. For the Fibonacci chain, Ref.~\cite{zachary2009hyperuniform} established
$\bar{\Lambda} \approx 0.201$ by an analytical theta-function method.

\subsection{Substitution Tilings and the Eigenvalue Formula}
\label{sec:substitution}

A one-dimensional \textit{substitution tiling} is an aperiodic sequence of tiles generated
by repeatedly applying a fixed rewriting rule to each letter of a finite alphabet. Starting
from a seed word, each letter is replaced by a fixed string of letters; iteration produces
an increasingly long sequence that, when each tile is assigned a physical length and the
tiles are stacked end to end, defines a one-dimensional point pattern. These tilings
provide some of the cleanest examples of quasicrystalline order: they are fully
deterministic, can be generated to arbitrary length by iteration, and their algebraic
properties encode their hyperuniformity class in a precise and computable way.

The growth of a substitution tiling is governed by its \textit{substitution matrix} $M$,
whose entry $M_{ij}$ counts how many tiles of type $j$ appear in the substituted image of
a single tile of type $i$. For the Fibonacci chain with alphabet $\{S, L\}$ and rules
$S \to L$, $L \to LS$, the substitution matrix is
\begin{equation}
  M = \begin{pmatrix} 0 & 1 \\ 1 & 1 \end{pmatrix},
  \label{eq:fibonacci-matrix}
\end{equation}
with eigenvalues $\lambda_1 = \tau = (1+\sqrt{5})/2 \approx 1.618$ (the golden ratio) and
$\lambda_2 = -1/\tau \approx -0.618$. The Perron-Frobenius theorem guarantees that the
largest eigenvalue $\lambda_1$ is positive and real, and its eigenvector gives the
asymptotic relative frequencies of each tile type. Tile lengths are assigned by requiring
that inflating every tile by $\lambda_1$ and then applying the substitution leaves the
total domain length invariant, which fixes the ratio $L/S = \lambda_1$. For the Fibonacci
chain this gives $L = \tau S$: the long tile is $\tau$ times the short tile.

The key result connecting the matrix spectrum to hyperuniformity is due to
Oğuz et al.~\cite{oguz2017hyperuniform,oguz2019substitution}: for a substitution tiling
with Perron-Frobenius eigenvalue $\lambda_1$ and second-largest-magnitude eigenvalue
$\lambda_2$,
\begin{equation}
  \alpha = 1 - \frac{2\ln|\lambda_2|}{\ln|\lambda_1|}.
  \label{eq:eigenvalue-formula}
\end{equation}
This formula is the central predictive tool of this work: $\alpha$ is determined entirely
by the substitution matrix spectrum, with no simulation required. The tiling is hyperuniform
when $\alpha > 0$, which requires $|\lambda_2| < |\lambda_1|^{1/2}$. The physical
interpretation is clear: $|\lambda_1|$ sets the inflation rate under substitution, while
$|\lambda_2|$ controls how quickly spatial correlations across tile boundaries decay. A
larger ratio $|\lambda_2|/|\lambda_1|^{1/2}$ means slower decorrelation and weaker
hyperuniformity.

For the metallic-mean chains and for the Bombieri-Taylor cubic substitution, the largest
eigenvalue $\lambda_1$ is a
\textit{Pisot-Vijayaraghavan (PV) number}: an algebraic integer greater than one all of
whose algebraic conjugates lie strictly inside the unit circle
($|\lambda_j| < 1$ for $j \geq 2$)~\cite{bombieri1986which,oguz2019substitution}.
The PV property guarantees a pure Bragg diffraction spectrum --- the tiling is a true
quasicrystal in the crystallographic sense, with delta-function peaks in $S(k)$ at a dense
set of wavevectors. Equation~\eqref{eq:eigenvalue-formula} also reveals the structural
reason why all $2\times 2$ metallic-mean matrices are locked to $\alpha = 3$ and why
accessing intermediate exponents requires a higher-rank substitution matrix.

%==============================================================================
\section{Methods}
\label{sec:methods}
%==============================================================================

We characterize each point pattern using three complementary observables: the number
variance $\sigma^2(R)$, the structure factor $S(k)$, and the diffusion
spreadability $\mathcal{S}(t)$~\cite{torquato2021spreadability,wang2022spreadability}.
These are computed from finite point patterns generated to sizes large enough that
finite-size effects are negligible on the window and wavenumber scales of interest.
This section describes how each pattern type is generated, how each observable is
computed, and how the results are validated against exact benchmarks.

\subsection{Pattern Generation}
\label{sec:generation}

\textbf{Substitution tilings.}
Each substitution tiling is generated by iterating the substitution rule on a seed word,
continuing until the total point count exceeds a target $N_\text{target}$. At each
iteration every tile of type $i$ is replaced by the string of tiles prescribed by row $i$
of the substitution matrix $M$; after $n$ iterations the total tile count grows
approximately as $\lambda_1^n$. Tile lengths are set by the Perron-Frobenius eigenvector
of $M$: if $v_i > 0$ is the $i$-th component, the length of tile type $i$ is
$\ell_i = v_i / \min_j v_j$, so that the shortest tile always has unit length. Points are
placed at tile boundaries. The resulting pattern has $N$ points on a domain of length
$L \approx N / \rho$, where the mean density is $\rho = 1/\bar\ell$ and
$\bar\ell = \sum_i f_i \ell_i$ is the frequency-weighted mean tile length computed from
the asymptotic tile frequencies $f_i$ given by the normalized Perron-Frobenius eigenvector.
For the metallic-mean chains, $N$ ranges from $1.5 \times 10^4$ to $\sim 4 \times 10^7$
depending on convergence requirements; for the Bombieri-Taylor cubic, 25 substitution
iterations yield $N \approx 9.5 \times 10^6$.

\textbf{Uncorrelated random lattices.}
The uncorrelated random lattice (URL) is generated by taking the integer lattice
$\{1, 2, \ldots, N\}$ and displacing each site independently by a random variate drawn
from a uniform distribution parameterized by amplitude $a$. Realizations at $a = 0.1$,
$0.5$, and $1.0$ are used; for each, the variance is computed from a single large
realization ($N = 10^6$) and compared with the exact formula $\bar\Lambda = (1+a^2)/6$
derived by Klatt, Kim, and Torquato~\cite{klatt2020perturbed}.

\textbf{Stealthy configurations.}
Stealthy hyperuniform configurations, for which $S(k) = 0$ for all $|k| < K^*$, were
generated by a collaborating group using collective coordinate
optimization~\cite{torquato2015ensemble}. The ensemble consists of approximately 4,316
independent configurations per value of the stealthiness ratio $\chi \in \{0.1, 0.2, 0.3\}$,
each containing $N = 2000$ points at mean density $\rho = 1$.
Rather than a single large system, the high-statistics ensemble average over thousands of
configurations suppresses sample-to-sample noise and provides reliable estimates of both
$S(k)$ and $\bar\Lambda$.

\subsection{Number Variance and Surface-Area Coefficient}
\label{sec:variance-method}

For a pattern of $N$ points on a domain of length $L$, the number variance $\sigma^2(R)$
at window half-width $R$ is estimated by Monte Carlo sampling. We place $N_\text{samp} =
30{,}000$ windows of half-width $R$ at positions drawn uniformly at random along the
domain, count the number of points $n_i$ in each window, and compute
$\sigma^2(R) = \langle n^2 \rangle - \langle n \rangle^2$. The window half-width is swept
over a logarithmic grid from $R_\text{min} = 1/\rho$ (one inter-point spacing) to
$R_\text{max} = L/4$; the upper bound ensures that every window placement is fully
contained within the domain and that the sampling is not dominated by boundary effects.
At each $R$, the estimator is essentially unbiased because the window placements are
independent of the point pattern itself.

The surface-area coefficient $\bar\Lambda$ is extracted as the running average of the
variance curve~\cite{torquato2003local},
\begin{equation}
  \bar\Lambda(R_\text{max}) = \frac{1}{R_\text{max}} \int_0^{R_\text{max}} \sigma^2(R')\,dR'.
  \label{eq:lambda-running}
\end{equation}
For Class~I patterns, this running average converges to a plateau as $R_\text{max}$
increases, providing the estimate of $\bar\Lambda$. Convergence is monitored by checking
that Eq.~\eqref{eq:lambda-running} varies by less than $1\%$ as $R_\text{max}$ is doubled;
for all substitution tilings and URL patterns studied here, this criterion is met well within
the available domain length. For stealthy configurations, $\bar\Lambda$ is computed from
the ensemble-averaged variance curve at $R_\text{max} = N/(4\rho)$, where $N = 2000$
provides sufficient range for convergence at all three $\chi$ values studied.

\subsection{Structure Factor and Hyperuniformity Exponent}
\label{sec:sfactor-method}

The structure factor $S(k)$ is evaluated on the discrete wavenumber grid $k_n = 2\pi n / L$
for integers $n = 1, 2, \ldots, N/2$ using the direct sum in Eq.~\eqref{eq:sk-def},
computed via a fast Fourier transform. The low-$k$ region, which determines the
hyperuniformity class and exponent, is sampled at fine wavenumber resolution and the
$S(k_n)$ values are logarithmically binned to reduce scatter before fitting.

For patterns with a smooth $S(k)$ near $k = 0$ (URL, stable distributions, stealthy),
the exponent $\alpha$ is extracted from an unweighted linear fit to $\log S(k)$ versus
$\log k$ over the wavenumber range where a clear power law is evident. For substitution
tilings, $S(k)$ consists entirely of isolated Bragg peaks rather than a smooth envelope;
a direct log-log fit to peak heights is unreliable. Instead, we follow Oğuz
et al.~\cite{oguz2019substitution} and fit the cumulative intensity
\begin{equation}
  Z(k) = 2\int_0^k S(q)\,dq \sim C\,k^{1+\alpha},
  \label{eq:Zk-def}
\end{equation}
which is a smooth, monotonically increasing function of $k$ even when $S(k)$ is a
discrete set of peaks. An unweighted linear fit to $\log Z(k)$ versus $\log k$ over a
range spanning at least one decade in $k$ yields $\alpha$ with uncertainty estimated from
the fit residuals. As an independent cross-check, $\alpha$ is also estimated from the
log-log slope of $\sigma^2(R)$ at large $R$; for all Class~I patterns studied, the two
methods agree to within $5\%$.

\subsection{Diffusion Spreadability}
\label{sec:spreadability-method}

The diffusion spreadability $\mathcal{S}(t)$ quantifies the fraction of space not yet
reached by a species diffusing from the point pattern at time
$t$~\cite{torquato2021spreadability,wang2022spreadability}. In one dimension and for a
two-phase medium with packing fraction $\phi_2$, the spreadability is related to $S(k)$ via
\begin{equation}
  \mathcal{S}(t) = \phi_2^2 + \frac{1}{\pi}\int_0^\infty S(k)\,e^{-2k^2 t}\,dk.
  \label{eq:spreadability-def}
\end{equation}
For Class~I and Class~II hyperuniform patterns, the excess spreadability
$E(t) = \mathcal{S}(t) - \phi_2^2$ decays as a power law at long times: $E(t) \sim
t^{-(1+\alpha)/2}$~\cite{torquato2021spreadability}, while stealthy configurations exhibit
exponential decay. We evaluate $\mathcal{S}(t)$ numerically by discretizing the integral
over the precomputed structure factor values $\{S(k_n)\}$ at a two-phase packing fraction
$\phi_2 = 0.35$~\cite{hitinbialus2024}. In practice, we use spreadability primarily as an
independent qualitative consistency check on the class assignment. For substitution tilings
with dense Bragg spectra --- in particular the Bombieri-Taylor cubic, which has 32,517
Bragg peaks in $k < 50$ --- the integral~\eqref{eq:spreadability-def} requires
extremely long times to converge, and the eigenvalue formula Eq.~\eqref{eq:eigenvalue-formula}
is used instead for precise $\alpha$ estimates.

\subsection{Validation Benchmarks}
\label{sec:validation}

Before applying these methods to the patterns under study, all three observables are
validated against exact analytical benchmarks.

\textbf{Poisson process.}
For a Poisson point process at density $\rho = 1$ and system size $N = 10^6$, the
exact results are $\sigma^2(R) = 2\rho R$ and $S(k) = 1$ for all $k$. Numerical estimates
reproduce the Poisson variance to within $1.1\%$ across three decades of window size and
$S(k) = 1.000 \pm 0.002$ at all sampled wavenumbers, confirming that the window-sampling
estimator and the FFT-based structure factor introduce no systematic bias.

\textbf{Uncorrelated random lattice.}
For the URL at amplitudes $a = 0.1$, $0.5$, and $1.0$, the exact surface-area coefficient
is $\bar\Lambda = (1+a^2)/6$, derived analytically in Ref.~\cite{klatt2020perturbed}.
Numerical estimates agree with this formula to within $0.3\%$ at all three amplitudes,
validating both the number-variance estimator and the running-average extraction of
$\bar\Lambda$.

\textbf{Fibonacci chain.}
For the Fibonacci quasicrystal, Zachary and Torquato~\cite{zachary2009hyperuniform}
computed $\bar\Lambda = 0.20110$ analytically using a theta-function representation of
the pair correlation function. Our Monte Carlo estimate at $N \approx 1.5 \times 10^7$
gives $\bar\Lambda = 0.200 \pm 0.001$, consistent with the analytical value at the level
of numerical precision; this provides a cross-check on the method for quasiperiodic
patterns where no simple closed-form formula exists.

%==============================================================================
% Bibliography
%==============================================================================
\printbibliography

\end{document}
